---
title: Drawing mathematical flowers with R
author: ~
date: '2019-08-01'
slug: drawing-mathematical-flowers-with-r
categories: []
tags: []
---

# Exploring visualisation with ggplot while playing with numbers
In this post we'll create a 'flowery' pattern using geom_point function. Let's see how we build toward it step by step. (PS- I did this exercise as a part of a project in datacamp)

```{r setup, include=FALSE}
      knitr::opts_chunk$set(collapse = TRUE)
      library(ggplot2)
      options(repr.plot.width = 4, repr.plot.height = 4)

```

#### Let's begin by creating a basic circle in R

```{r message = FALSE, warning = FALSE}
t <- seq(0, 2*pi, length.out = 50)
x <- sin(t)
y <- cos(t)
df <- data.frame(t, x, y)

# Make a scatter plot of points in a circle
p <- ggplot(df, aes(x, y))
p + geom_point()
```

#### Let us now increase the points, add the magic of golden ratio, and create a spiralling effect

```{r message = FALSE, warning = FALSE}
# Defining the number of points
points <- 500
# Defining the Golden Angle
angle <- pi*(3-sqrt(5))

t <- (1:points) * angle
x <- sin(t)
y <-cos(t)
df <- data.frame(t, x, y)

# Make a scatter plot of points in a spiral
p <- ggplot(df, aes(x*t, y*t))
p + geom_point()
```

#### Next we move on to removing the distractions and adding color

```{r message = FALSE, warning = FALSE}
p + geom_point(size = 8, alpha = 0.5, color = "darkgreen") + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_rect(fill="white"),panel.border=element_blank(),panel.grid = element_blank(), panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_rect(fill="white"))
```

#### Next we play more with colour and shapes of points and attempt to create a sunflower like image using a magenta background and yellow points

```{r message = FALSE, warning = FALSE}
p + geom_point(aes(size = t), alpha = 0.5, color = "yellow", shape = 17) + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_rect(fill="darkmagenta"))
```

#### Lastly, with another experimentation, let's add a hynotic feel to the flower image by increasing the points, increasing the size of the points, and altering the angle

```{r message = FALSE, warning = FALSE}
angle <- 13*pi/180
points <- 2000

t <- (1:points)*angle
x <- sin(t)
y <- cos(t)

df <- data.frame(t, x, y)

p <- ggplot(df, aes(x*t, y*t))
p + geom_point(size = 80, alpha = 0.1, color = "magenta4", shape = 1) + theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_rect(fill="white"),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_rect(fill="white"))
```


